---
title: "ANTI-FUNGAL TREATMENT RESULTS IN LITTLE TO NO SHORT-TERM SIDE EFFECTS ON HOST PERFORMANCE, FITNESS, AND BEHAVIOR"
author: "G. V. DiRenzo"
date: "5/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load libraries
library(lme4)
```


# Reading in and formatting the data

```{r}
# set working directory
setwd("~/Dropbox/Active manuscripts/Needs Work/Behavior/")

# Read in the jumping data
dat1 <- read.csv("./Data/Jumping/distancejumps_GVD_23_March_2018.csv")

# Look at structure of data
  # 398 observations of 12 variables
str(dat1)
  # $ID_Number = individual id number of the frog
  # $Distance_1_mm = distance of 1st jump in mm
  # $Distance_2_mm = distance of 2nd jump in mm
  # $Total = sum of 1st and 2nd jump
  # $Day = Day of jumping trial
  # $Mo = month (numerical) of jumping trial
  # $Month = month (name) of jumping trial
  # $Year = year of jumping trial
  # $Average_Total_Distance_mm = average distance jumped from 2 consecutive jumps
  # $Exp_Day = experimental day of jumping trial
  # $Jump_time = when the jumping trial took place - first, middle, or last 
  # $Bd_load = Bd load of amphibian during jumping trial

# Convert both of the Distance columns to numeric
dat1$Distance_1_mm <- as.numeric(as.character(dat1$Distance_1_mm))
dat1$Distance_2_mm <- as.numeric(as.character(dat1$Distance_2_mm))

# read in treatment groups
dat2 <- read.csv("./Data/Treatments/FINAL TREATMENTS/treatments_long_format.csv")

# Treatment information for frogs 
str(dat2)
# 115 obs. of  5 variables:
 #$ ID_Number = individual id number of the frog
 #$ Treatment = Treatment
 #$ Infection = Infected or Uninfected
 #$ New_treatment = Summary treatment
 #$ Group = Not applicable


# Merge the 2 data files by ID_Number
dat <- merge(dat1, dat2, by= "ID_Number")

# Look at the new data frame
str(dat)

# Convert ID Number to a factor (like names)- it is not a continuous variable
dat$ID_Number <- as.factor(dat$ID_Number)

# Average jump between the 2 jumps
dat$Two_jump_average <- apply(cbind(dat$Distance_1_mm, dat$Distance_2_mm), 1, mean, na.rm = TRUE)

# Remove any treatment with the work "Pre" in it
dat <- dat[-c(grep("Pre-", dat$Treatment)),]

# Drop the levels
dat <- droplevels(dat)

# Look at data structure
str(dat)

# Create the table of sample sizes
library(plyr)
ddply(.data = dat, .variable = "Treatment", .fun = summarize,
      sample_size = length(unique(ID_Number)))
```

#  Appendix S1. Host performance

## (1) LMM: total jumping distance ~ treatment*experimental day + (1| ID number)

```{r}
# We will be analyzing the jumping trials
# We will compare the pre-jumping distance (i.e., pre infection, pre antifungal application) to the post-jumping distance (i.e., post infection, post antifungal application)

# Remove experimental days 26 and 40 (We conducted jumping trials on these days that are not analyzed here)
jump <- dat[-c(dat$Exp_Day == 26 | dat$Exp_Day == 40),]

# Create an empty column
jump$time <- NA

# If experimentdal day is less than 40- give it the name "Pre", else give it the name "Post"
jump$time <- ifelse(jump$Exp_Day < 40, "Pre", "Post")

# Make the column into a factor with 2 levels
jump$time <- factor(jump$time, levels = c("Pre", "Post"))

# Analyze the total jumping distance (response variable) with time * treatment interaction as fixed effects and individual ID as the random effect
mod <- lmer(Total ~ time*Treatment + (1|ID_Number), data = jump)

# Look at model summary
summary(mod)

#---- Post-hoc analysis- Look at pairwise differences
# Load the library
library("emmeans")

# Visualize the interaction
# Appendix S1 (B)
emmip(mod, Treatment ~ time, ylab = "Total jumping distance (mm)", xlab = "Time")

# save the plot
library(ggplot2)
ggsave("jumping.pdf", height = 5, width = 5)

# Compare the total jumping distance between "Pre" and "Post" for each treatment
post_hoc <- emmeans(mod, "time", by = "Treatment")

# Calculate p-values for each pair
pairs(post_hoc)
```

```{r}

####-------- Make the plot: Figure 2
# Load the libraries
library(ggplot2)
library(plyr)

mat4 <- ddply(.data = jump, .variable = c("Treatment", "time"), .fun = summarize, 
              mean = mean(Total), 
              upper = mean(Total) +(sd(Total)/sqrt(length(Total))),
              lower = mean(Total) -(sd(Total)/sqrt(length(Total))))

ggplot(data = mat4, aes(y = mean, x = Treatment, fill = time)) + 
    geom_bar(stat="identity", color="black", position=position_dodge()) +
  scale_fill_manual(values = c("gray47", "goldenrod"))+
    geom_errorbar(aes(ymin=lower, ymax=upper), width=.2, position=position_dodge(.9)) +
    theme_bw()+ 
  ylab("Total distance jumped (mm)")+
  xlab("Treatment")+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.x=element_text(size=12))+
  labs(fill = "Time") +
  annotate("text", x = 3, y = 200, label = "*", size = 10)

ggsave("Jumping bar.pdf", height = 5, width = 8)
```

## (II) LMM: total jumping distance ~ log10(Bd load + 1)* treatment + experimental day + (1| ID number)
 
```{r}

# Read in the jumping data
dat1 <- read.csv("~/Dropbox/Active manuscripts/Needs Work/Behavior/Data/Jumping/distancejumps_GVD_23_March_2018.csv")

# Look at the data structure
str(dat1)
# Similar as above - # 398 obs.

# Read in treatment groups
dat2 <- read.csv("~/Dropbox/Active manuscripts/Needs Work/Behavior/Data/Treatments/FINAL TREATMENTS/treatments_long_format.csv")

# look at data structure
str(dat2)
# Similar as above - # 115 obs.

# Merge the 2 data files
dat <- merge(dat1, dat2, all = TRUE)

# Look at data structure
str(dat)
# 398 obs

# Remove pre group- Not used for this paper
dat <- dat[-c(grep("Pre-", dat$Treatment)),]
# 324 obs.

# Drop any unsed levels in the factor
dat <- droplevels(dat)

# Look at the data structure 
str(dat)
# 324 obs

# Keep only infected groups
# dat <- dat[dat$Infection == "Infected",]

# Convert ID Number to a factor (like names)- it is not a continuous variable
dat$ID_Number <- as.factor(dat$ID_Number)

# Remove any individuals with no Bd load
dat <- dat[is.na(dat$Bd_load) == FALSE,]

# Drop any unused levels
dat <- droplevels(dat)

# Look at data structure
str(dat)
# 79 obs

# Make a new column with Bd load log transformed
dat$log10Bd <- log10(dat$Bd+1)

# load libraries
library(lme4)
library(nlme)
library("emmeans")

# Used generalized linear mixed effect model with log10(Bd+1) load as the response variable and 
mod <- lme(Total ~  log10Bd * Treatment + Exp_Day, random = ~1|ID_Number, data = dat)

# Look at summary
summary(mod)
  # log10Bd is not significant
  # Next, we will test to see if the slopes for each treatment are different
  # And we will test to see if the intercepts are different among treatments


#Create a new data frame
dat2 <- expand.grid(Exp_Day = 30, #unique(dat2$Exp_Day.y),
                    Treatment = levels(dat$Treatment),
                    log10Bd = seq(from = 0, to = 6, by= 1),
                    ID_Number = 1 #unique(dat2$ID_Number)
                    )

# Call library
library(AICcmodavg)
# mean model fit
dat2$pred_mean <- predictSE(mod = mod, newdata= dat2, type = "response", backTransform = TRUE)$fit
# Upper
dat2$pred_upper <- predictSE(mod = mod, newdata= dat2, type = "response", backTransform = TRUE)$fit +
                   predictSE(mod = mod, newdata= dat2, type = "response", backTransform = TRUE)$se.fit
# Lower
dat2$pred_lower <- predictSE(mod = mod, newdata= dat2, type = "response", backTransform = TRUE)$fit -
                   predictSE(mod = mod, newdata= dat2, type = "response", backTransform = TRUE)$se.fit

dat2$ID_Number <- as.factor(dat2$ID_Number)

# Load the library
library(ggplot2)

# Make a plot
  # Total jumping distance = y-axis
  # log10 Bd load = x-axis
  # Treatments = colors
ggplot(data = dat, aes(y = Total, x = log10Bd, col = Treatment))+ 
  geom_jitter(height = 0.01, width = 0.1) + 
 geom_ribbon(data = dat2, aes(x = log10Bd,
                              y = pred_mean,
                              ymin= pred_lower, 
                              ymax = pred_upper, fill = Treatment), alpha = 0.3, colour = NA)+
  geom_line(data = dat2, aes(x = log10Bd, y = pred_mean, col = Treatment), size = 1.5)+
  scale_color_manual(values = c("black", "deepskyblue3", "goldenrod3"))+
  scale_fill_manual(values = c("black", "deepskyblue3", "goldenrod3"))+
  ylab("Total jumping distance (mm)") + 
  xlab(expression(paste(log[10], "(", italic(Bd), " load + 1)")))+ 
  theme_bw()+
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    axis.text.y=element_text(size=14),
    axis.title.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    axis.title.x=element_text(size=14))
  
ggsave("Jumping distance v Bd load.pdf", height = 5, width = 7)


# Interaction plot
emmip(mod, Treatment ~ log10Bd, cov.reduce = range,
      xlab = "log10(Bd load + 1)", ylab = "Total jumping distance (mm)") +
  scale_x_discrete(labels = c(0, 6))

ggsave("Jumping distance Bd load interaction plot.pdf", height = 5, width = 7)
  

#---- Look at pairwise differences in slope estimates among treatments
lst <- emtrends(mod, ~ Treatment, var = "log10Bd")
multcomp::cld(lst)
# Interaction is not significant


# Compare the intercept for total jumping distance between treatments
post_hoc <- emmeans(mod, "Treatment")

# Calculate p-values for each pair
pairs(post_hoc)
# Intercept not significantly different between pairs of treatments

```


# Appendix S2. Host fitness 

## (I) LMM: weight ~ treatment*experimental day + (1| ID number)

```{r}
# Influence of treatment on growth
growth <- read.csv("~/Dropbox/Active manuscripts/Needs Work/Behavior/Data/Arrival_day/Arrival_day_measurements.csv")

# Look at data
str(growth)
#   # 'data.frame':	874 obs. of  13 variables:
#  $ Month = Month of measurement
#  $ Day  = Day of measurement
#  $ Year = Year of measurement
#  $ ID_Number  = ID number of individual
#  $ Weight_g  = weight in grams of individual
#  $ SVL_mm  = snout to vent length of individual
#  $ Swab_ID  = Swab ID of sample collected
#  $ Sprayed_YN  = Was the individual sprayed with water?
#  $ Fell_On_Floor_YN  = Did the individual fall on the floor?
#  $ Feed_YN  = Was the individual fed?
#  $ Amount_of_fruit_flies_fed = The number of fruit flies added to the cage
#  $ Notes = any other notes
#  $ Exp_Day = experimental day

# Download treatment data
treatments <- read.csv("~/Dropbox/Active manuscripts/Needs Work/Behavior/Data/Treatments/FINAL TREATMENTS/treatments_long_format.csv")

# Same file as before with the treatment information

# Convert ID number to factor- used as a name- not continuous variable
treatments$ID_Number <- as.factor(treatments$ID_Number)

# Merge the matrix with the treatments
mat <- merge(growth, treatments, all = TRUE)

# Look at the merged file structure
str(mat)

# Remove any treatments with the word "pre"
mat <- mat[-c(grep("Pre-", mat$Treatment)),]

# Drop any unused levels
mat <- droplevels(mat)

# Convert ID number to factor- used as a name- not continuous variable
mat$ID_Number <- as.factor(mat$ID_Number)

## Analyze the frog's weight (response variable) with experimental day * treatment interaction as fixed effects and individual ID as the random effect
library(lme4)
mod <- lmer(Weight_g ~ Treatment*Exp_Day + (1|ID_Number), data = mat)

# Look at summary of the model
summary(mod)
```

```{r}

#---- Look at pairwise differences
# load the library
library("emmeans")


# Visualize the interaction
# Appendix S2 (B)
emmip(mod, Treatment ~ Exp_Day, ylab = "Weight (g)", xlab = "Experimental day", cov.reduce = FALSE)

# save the plot
library(ggplot2)
ggsave("weight- emmip.pdf", height = 5, width = 8)


# Here, we are interested in determining if the slope (change in weight over time estimate) differs among treatments
lst <- emtrends(mod, ~ Treatment, var = "Exp_Day")

multcomp::cld(lst)
```


```{r}

mat <- mat[is.na(mat$Treatment) == FALSE,]

# Drop unused levels
mat <- droplevels(mat)

# Put the treatments in the order you want them to appear in the plot
mat$Treatment <- factor(mat$Treatment, levels = c("Bd", "Bd+BMP-NTf2", "Bd+ITCZ", "BMP-NTf2", "Control", "ITCZ"))

#
ann_text <- data.frame(Weight_g = rep(1, times = 6),
                       Exp_Day  = rep(5, times = 6),
                       ID_Number = rep(1, times = 6),
                       lab = c(1, 1, 2, 3, 3, 3),
                       Treatment = factor(c("Bd", "Bd+BMP-NTf2", "Bd+ITCZ", "BMP-NTf2", "Control", "ITCZ"), 
                                          levels = c("Bd", "Bd+BMP-NTf2", "Bd+ITCZ", "BMP-NTf2", "Control", "ITCZ")))

# Make plot
ggplot(data = mat, aes(y = Weight_g, x = Exp_Day, col = as.factor(ID_Number))) + 
  geom_point()+
  geom_line()+
    geom_smooth(data = mat, aes(x = Exp_Day, y  = Weight_g, col = Treatment), method = "lm", col = "black")+
  facet_wrap(~Treatment)+
    theme_bw()+
  ylab("Mass (g)")+
  xlab("Experimental day")+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=12),
        axis.text.x=element_text(size=12),
        legend.position = "none",
        axis.title.x=element_text(size=12),
        strip.background = element_rect("white"))+
  geom_text(data = ann_text, label = c(1, 1, 2, 3, 3, 3), col = "black")

ggsave("Weight through time.pdf", height = 5, width = 7)

```

## (II) LMM: weight ~ log10(Bd load + 1)* treatment + experimental day + (1| ID number)

```{r}
# Here, we are determining the effect of pathogen load on weight
growth <- read.csv("~/Dropbox/Active manuscripts/Needs Work/Behavior/Data/Arrival_day/Arrival_day_measurements.csv")

# Look at data structure
str(growth)
# Same file as before with the growth information

# Download treatment data
treatments <- read.csv("~/Dropbox/Active manuscripts/Needs Work/Behavior/Data/Treatments/FINAL TREATMENTS/treatments_long_format.csv")
# Same file as before with the treatment information

# Convert ID number to factor
treatments$ID_Number <- as.factor(treatments$ID_Number)

# Merge the matrix with the treatments
mat <- merge(growth, treatments, all = TRUE)

# Look at new data frame structure
str(mat)

# Remove any treatments with "Pre-" name
mat <- mat[-c(grep("Pre-", mat$Treatment)),]

# Drop any levels not used in the data frame
mat <- droplevels(mat)

# Convert ID number to factor
mat$ID_Number <- as.factor(mat$ID_Number)

# Read in the infection data
Bd <- read.csv("~/Dropbox/Active manuscripts/Needs Work/Behavior/Data/Pathogen_load/Bd-DT2.csv")

# Look at the data structure
str(Bd)
#'data.frame':	439 obs. of  6 variables:
# $ Bd  = zoospore genomic equivalents detected on the swabs
# $ day  = day sample collected
# $ Month  = month sampled collected
# $ year   = year sample collected
# $ Exp_Day = experimental day sample collected
# $ ID_Number = individual ID

# Remove any rows without Bd infection information
Bd <- Bd[is.na(Bd$Bd) == FALSE,]

# Rename the second column to "Day"
colnames(Bd)[2] <- "Day"

# Rename the fourth column to "Year"
colnames(Bd)[4] <- "Year"

# Convert ID number to factor
Bd$ID_Number <- as.factor(Bd$ID_Number)

# Merge the two dataframes together along all columns that match between them
dat9 <- merge(x = Bd, y = mat, by = c("Day", "Month", "Year", "ID_Number"))

# Remove any rows without Bd information
dat9 <- dat9[is.na(dat9$Bd) == FALSE,]

# Look at data structure
str(dat9)

# Drop any levels not used
dat9 <- droplevels(dat9)

# Remove rows without treatment information
mat2 <- dat9[is.na(dat9$Treatment) == FALSE,]

# Drop unused levels
mat2 <- droplevels(mat2)

# Make a new column with Bd load log transformed
mat2$log10Bd <- log10(mat2$Bd+1)
  
# load libraries
library(lme4)
library(nlme)
library("emmeans")

# Used generalized linear mixed effect model with log10(Bd+1) load as the response variable and 
mod <- lme(Weight_g ~  log10Bd * Treatment + Exp_Day.y, random = ~1|ID_Number, data = mat2)

summary(mod)

#Create a new data frame
mat3 <- expand.grid(Exp_Day.y = 30, #unique(mat2$Exp_Day.y),
                    Treatment = levels(mat2$Treatment),
                    log10Bd = seq(from = 0, to = 6, by= 1),
                    ID_Number = 1 #unique(mat2$ID_Number)
                    )

# Call library
library(AICcmodavg)
# mean model fit
mat3$pred_mean <- predictSE(mod = mod, newdata= mat3, type = "response", backTransform = TRUE)$fit
# Upper
mat3$pred_upper <- predictSE(mod = mod, newdata= mat3, type = "response", backTransform = TRUE)$fit +
                   predictSE(mod = mod, newdata= mat3, type = "response", backTransform = TRUE)$se.fit
# Lower
mat3$pred_lower <- predictSE(mod = mod, newdata= mat3, type = "response", backTransform = TRUE)$fit -
                   predictSE(mod = mod, newdata= mat3, type = "response", backTransform = TRUE)$se.fit

mat3$ID_Number <- as.factor(mat3$ID_Number)

# Plot data
ggplot(data = mat2, aes(y = Weight_g, x = log10Bd, col = Treatment))+ 
  geom_jitter(height = 0.01, width = 0.1) + 
 geom_ribbon(data = mat3, aes(x = log10Bd,
                              y = pred_mean,
                              ymin= pred_lower, 
                              ymax = pred_upper, fill = Treatment), alpha = 0.3, colour = NA)+
  geom_line(data = mat3, aes(x = log10Bd, y = pred_mean, col = Treatment), size = 1.5)+
  scale_color_manual(values = c("black", "deepskyblue3", "goldenrod3"))+
  scale_fill_manual(values = c("black", "deepskyblue3", "goldenrod3"))+
  ylab("Weight (g)") + 
  xlab(expression(paste(log[10], "(", italic(Bd), " load + 1)")))+ 
  theme_bw()+
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    axis.text.y=element_text(size=14),
    axis.title.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    axis.title.x=element_text(size=14))

ggsave("Weight v Bd load.pdf", height = 5, width = 7)

# Interaction plot
emmip(mod, Treatment ~ log10Bd, cov.reduce = range,
      xlab = "log10(Bd load + 1)", ylab = "Weight (g)") +
  scale_x_discrete(labels = c(0, 6))

ggsave("Weight Bd load interaction plot.pdf", height = 5, width = 7)
  

#---- Look at pairwise differences in slope estimates among treatments
lst <- emtrends(mod, ~ Treatment, var = "log10Bd")
multcomp::cld(lst)


# Compare the intercept for total jumping distance between treatments
post_hoc <- emmeans(mod, "Treatment")

# Calculate p-values for each pair
pairs(post_hoc)
# Intercept not significantly different between pairs of treatments


```

# Appendix S3. Evaluating host behavior: 

```{r}
# library
require(foreign)
require(nnet)
require(ggplot2)
require(reshape2)
library("car")

# Load and format the data

# Read data
dat <- read.csv("~/Dropbox/Active manuscripts/Needs Work/Behavior/Data/Habitat/habitatworksheet.csv")

# Look at data

str(dat)
#'data.frame':	2906 obs. of  10 variables:
# $ Day = day of observation 
# $ Month = month of observation
# $ Year  = year of observation
# $ Frog  = Frog Id number
# $ Habitat.Notes = Frog habitat use
# $ Posture  = Frog posture
# $ Movement   = notes on movement/activity
# $ Time.Habitat = time of habtiat observation 
# $ Time.Posture = time of posture observation
# $ Notes   = other notes

# Look at the levels in habitat notes
unique(dat$Habitat.Notes)

# Calculate the number of times each observation occurred
table(dat$Habitat.Notes)

# There are clearly some levels that need to be grouped together
# Here we will use regular expressions to fix spelling mistakes

dat$Habitat.Notes <- gsub("against side  ", "ON_TOP_OR_ON_WALL", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("CROUCHED_BEHIND_HIDE ", "CROUCHED_BEHIND_HIDE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("CROUCHED_BEHIND_HIDE, halfway up on hide (torso)", "CROUCHED_BEHIND_HIDE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("CROUCHED_BEHIND_HIDE, halfway up ON_TOP_OR_ON_WALL(torso)", "CROUCHED_BEHIND_HIDE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("CROUCHED_BEHIND_HIDE, halfway up on hide (torso)", "CROUCHED_BEHIND_HIDE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("CROUCHED_BEHIND_HIDE)", "CROUCHED_BEHIND_HIDE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("in INSIDE_HIDE", "INSIDE_HIDE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("INSIDE_HIDE ", "INSIDE_HIDE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("leaning halway up on hide", "INSIDE_HIDE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("on ON_TOP_OR_ON_WALL", "ON_TOP_OR_ON_WALL", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("ON_SUBSTRATE ", "ON_SUBSTRATE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("ON_SUBSTRATE of cage ", "ON_SUBSTRATE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("ON_SUBSTRATEof cage ", "ON_SUBSTRATE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("ON_SUBSTRATEON_SUBSTRATE", "ON_SUBSTRATE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("ON_SUBSTRATEupright", "ON_SUBSTRATE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("ON_SUBSTRATE upright", "ON_SUBSTRATE", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("ON_TOP_OR_ON_WALL ", "ON_TOP_OR_ON_WALL", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("ON_TOP_OR_ON_WALL  ", "ON_TOP_OR_ON_WALL", dat$Habitat.Notes)

dat$Habitat.Notes <- gsub("side  ", "ON_TOP_OR_ON_WALL", dat$Habitat.Notes)

table(dat$Habitat.Notes)

# Remove levels with 1 observation
dat <- dat[dat$Habitat.Notes != "mid tank ",]

# remove rows without any observations
dat <- dat[is.na(dat$Habitat.Notes) == FALSE,]

# Check the structure of the data
str(dat)

# Convert Habitat.Notes to a factor
dat$Habitat.Notes <- as.factor(dat$Habitat.Notes)

# Check the number of levels
levels(dat$Habitat.Notes)

# Replace the name of the 4th column 
colnames(dat)[4] <- "ID_Number"

# Download treatment data
treatments <- read.csv("~/Dropbox/Active manuscripts/Needs Work/Behavior/Data/Treatments/FINAL TREATMENTS/treatments_long_format.csv")

# Download data
dates <- read.csv("~/Dropbox/Active manuscripts/Needs Work/Behavior/exp_dates.csv")

# Paste the day, month, and year
dates$dates <- paste(dates$Day, dates$Month, dates$Year, sep = "-")

# Look at dates
str(dates)

# Merge the matrix with the treatments
mat <- merge(dat, treatments, by = "ID_Number")

# Paste the day, month, and year together
mat$dates <- paste(mat$Day, mat$Month, mat$Year, sep = "-")

# Merge mat and dates dataframes
mat2 <- merge(mat, dates, all = TRUE)

# Remove any "pre-" treatment animals
mat2 <- mat2[-c(grep("Pre-", mat2$Treatment)),]

# Drop any unused levels
mat2 <- droplevels(mat2)

# Remove individuals without a treatment
mat2 <- mat2[is.na(mat2$Treatment) == FALSE,]

# Look at the head of the data frame
head(mat2)

# Create a new column
mat2$stage <- rep(0, times = nrow(mat2))

# Specify during what stage of the experiment the observation occurred
  # Experimental day < 20 = "before"
  # Experimental day == 20 = "during"
  # Experimental day > 20  & < 30 = "during"
  # Experimental day == 30 = "during"
  # Experimental day > 30  = "after"

for(i in 1:nrow(mat2)){
  if(mat2$Exp_Day[i] < 20){mat2$stage[i] <- "before"}
  if(mat2$Exp_Day[i] == 20){mat2$stage[i] <- "during"}
  if(mat2$Exp_Day[i] > 20 & mat2$Exp_Day[i] < 30) {mat2$stage[i] <- "during"}
  if(mat2$Exp_Da[i] == 30){mat2$stage[i] <- "during"}
  if(mat2$Exp_Day[i] > 30){mat2$stage[i] <- "after"}
}  

# Make this new stage column a factor
mat2$stage <- as.factor(mat2$stage)

# Look at the structure
str(mat2)

# Make a table of the number of observations by habitat choice and tratment
with(mat2, table (Habitat.Notes, Treatment))

# Fix the redundant level
mat2$Habitat.Notes <- gsub("ON_TOP_OR_ON_WALL ", "ON_TOP_OR_ON_WALL", mat2$Habitat.Notes)

# Change individual id numbers to a factor
mat2$ID_Number <- as.factor(mat2$ID_Number)

# Make a new column is Bd+ or Bd-
mat2$Group <- NA

# If Bd is in the treatment- then make the Group column = "Bd+"
mat2[grep("Bd", mat2$Treatment),]$Group <- "Bd+"

# If Bd is not in the treatment- then make the Group column = "Bd-"
mat2$Group <- ifelse(is.na(mat2$Group) == TRUE, "Bd-", mat2$Group)

# Make Group a factor 
mat2$Group <- as.factor(mat2$Group)

# Look at mat2 structure
str(mat2)

# Not used in manuscript but tested
test <- multinom(Habitat.Notes ~ Exp_Day * Group + ID_Number , data = mat2)

# Calculates type-II or type-III analysis-of-variance tables for multinomial model 
  # for multinomial logit and proportional-odds logit models, likelihood-ratio tests are calculated.
Anova(test, type = "III")
Anova(test, type = "II")

# Use a multinomial logit model to test the change in habitat use by treatment
test <- multinom(Habitat.Notes ~ Exp_Day * Treatment + ID_Number , data = mat2)

# Look at output- Appendix S3 A
summary(test)

# Calculates type-II or type-III analysis-of-variance tables for multinomial model 
  # for multinomial logit and proportional-odds logit models, likelihood-ratio tests are calculated.
Anova(test, type = "III")
Anova(test, type = "II")

# Make a plot to visualize the relationships

# Create a new data frame with covariates to 
dses <- data.frame(Treatment = rep(levels(mat2$Treatment), 
                                   times = length(unique(mat2$Exp_Day)) * 3),
                  Exp_Day = rep(unique(mat2$Exp_Day), 
                                each = length(levels(mat2$Treatment)) * 3),
                  ID_Number = rep(1, 
                                  each = length(levels(mat2$Treatment)) *length(unique(mat2$Exp_Day)) *3),
                  stage = rep(c("before", "during", "after"), 
                              each = length(levels(mat2$Treatment)) *
                                length(unique(mat2$Exp_Day))))

# Make the stage a factor
dses$stage <- as.factor(dses$stage)

# Make Id number a factor
dses$ID_Number <- as.factor(dses$ID_Number)

## store the predicted probabilities for each value of ses and write
pp.write <- cbind(dses, predict(test, dses, type = "probs"))

## melt data set to long form for plotting using ggplot2
lpp <- melt(pp.write, id.vars = c("Treatment", "stage", "ID_Number", "Exp_Day"), value.name = "probability")
head(lpp)

# Replace the names with other words
lpp$variable <- gsub("ON_SUBSTRATE", "On substrate", lpp$variable)

lpp$variable <- gsub("CROUCHED_BEHIND_HIDE", "Behind hide", lpp$variable)

lpp$variable <- gsub("INSIDE_HIDE", "Inside hide", lpp$variable)

lpp$variable <- gsub("ON_TOP_OR_ON_WALL", "On top or on wall", lpp$variable)

###--- Take average across individuals
library(plyr)

Summ <- ddply(.data = lpp, .variable = c("Treatment", "Exp_Day", "variable"), .fun = summarize,
      probability = mean(probability))

#--- Make a plot
cbbPalette <- c("#000000", "#009E73", "#e79f00", "#9ad0f3", "#0072B2", "#D55E00", "#CC79A7", "#F0E442", "#56B4E9")

# Create data frames for adding text to plot
# Before
ann_text1 <- data.frame(Exp_Day = rep(15, times = 4),
                       probability = rep(0.6, times = 4),
                       variable = c("Behind hide", 
                                     "Inside hide", 
                                     "On top or on wall", 
                                     "On substrate"),
                       Treatment = "Bd")

# During
ann_text2 <- data.frame(Exp_Day = rep(25, times = 4),
                       probability = rep(0.6, times = 4),
                       variable = c("Behind hide", 
                                     "Inside hide", 
                                     "On top or on wall", 
                                     "On substrate"),
                       Treatment = "Bd")

# After
ann_text3 <- data.frame(Exp_Day = rep(32, times = 4),
                       probability = rep(0.6, times = 4),
                       variable = c("Behind hide", 
                                     "Inside hide", 
                                     "On top or on wall", 
                                     "On substrate"),
                       Treatment = "Bd")

ggplot(Summ, aes(x = Exp_Day, y = probability, colour = Treatment)) + geom_line(size = 1.2) + facet_grid(variable ~., scales = "free")+ 
  scale_color_manual(values = cbbPalette)+
  ylab("Probability of habitat use") + xlab("Experimental Day") +
  geom_vline(xintercept = 20, col = "red", lty = 2)+
  geom_vline(xintercept = 30, col = "red", lty = 2)+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    strip.background = element_rect("white"),
    axis.text.y=element_text(size=14),
    axis.title.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    axis.title.x=element_text(size=14))+ 
  geom_text(data = ann_text1, show.legend = FALSE, label = c("Before", "", "", ""))+
  geom_text(data = ann_text2, show.legend = FALSE, label = c("During", "", "", ""))+
  geom_text(data = ann_text3, show.legend = FALSE, label = c("After", "", "", ""))
  
ggsave("Habitat choice v time.pdf", height = 8, width = 7)

```

